---
title: "county_level_indicators"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(readxl)
library(tidyverse)
library(reshape2)
library(skimr)
```

# All data loading for county-level indicators. Need to run in order!

```{r}
#create funtion to clean different county names formats

#Virginia has both counties and independent cities
va_independent_cities <- c("Alexandria", "Bristol", "Buena Vista", "Charlottesville", "Chesapeake", "Colonial Heights", "Covington", "Danville", "Emporia", "Fairfax City", "Falls Church", "Franklin", "Fredericksburg", "Galax", "Hampton", "Harrisonburg", "Hopewell", "Lexington", "Lynchburg", "Manassas", "Manassas Park", "Martinsville", "Newport News", "Norfolk", "Norton", "Petersburg", "Poquoson", "Portsmouth", "Radford", "Richmond", "Roanoke", "Salem", "Staunton", "Suffolk", "Virginia Beach", "Waynesboro", "Williamsburg", "Winchester")

#some independent cities have the same name as nearby counties
va_city_county_overlap <- c("Charles City", "James City", "Franklin", "Richmond", "Roanoke")

clean_county_names <- function(county_level_data, county){
  county_level_data <- county_level_data %>%
    mutate(county = tolower(as.character(county)))%>%
    mutate(county = tools::toTitleCase(as.character(county)))%>%
    mutate(county = ifelse(county %in% va_city_county_overlap, str_c(as.character(county), "County", sep = " "),
                           ifelse(county == "City of Suffolk", "Suffolk City", 
                                  ifelse(str_detect(county, "City") | str_detect(county, "County"), as.character(county),
                                         ifelse(county %in% va_independent_cities, str_c(as.character(county), "City", sep = " "), str_c(as.character(county), "County", sep = " "))))))
  
  return(county_level_data)
  
}

```


## ACS Data
Explanation of factors:

fips = FIPS Code
number_enrolled = population 3 years and over enrolled in school
number_enrolled_prek = population >3 y.o. enrolled in nursery school/preschool
number_enrolled_k = population enrolled in kindergarten
number_enrolled_elementary = population enrolled in grades 1-8
number_enrolled_hs = population enrolled in grades 9-12
number_enrolled_college = population enrolled in college or graduate school
pct_under_9th = percent of adults 25 and older with less than 9th grade education
pct_9th_12th = percent of adults >=25 with 9th-12th grade education, no diploma
pct_attained_hs = percent of adults with a high school diploma or equivalent, no higher
pct_some_college = percent of adults with some college education, no degree
pct_associates = precent of adults with associates degree, no higher
pct_bachelors = percent of adults with a bachelor's degree, no higher
pct_grad_degree = percent of adults with a graduate or professional degree
pct_hs_or_higher = percent of adults that are a high school graduate or higher
pct_college_or_higher = percent of adults with a bachelor's degree or higher

pct_owner_occupied_homes = percent of total housing units that are owner-occupied
pct_renter_occupied_homes = percent of total housing unites that are renter-occupied
pct_moved_after_2014 = percent of occupied housing units w residents who moved in 2015 or later
pct_moved_2010_2014 = percent with residents who moved in beween 2010 and 2014
pct_moved_2000_2009 = percent with residents who moved in beween 2000 and 2009
pct_moved_1990_1999 = percent with residents who moved in beween 1990 and 1999
pct_moved_1980_1989 = percent with residents who moved in beween 1980 and 1989
pct_moved_1979_earlier = percent with residents who moved in 1979 or earlier

pct_unemployed = unemployment rate
pct_commute_drive = percent of workers 16 and over who drive to work alone in a car, truck, or van
pct_commute_carpool = percent of workers who carpool to work
pct_commute_public = percent of workers who take public transportation (excluding taxi)
pct_commute_walk = percent of workers who walk
pct_commute_other = percent of workers who commute via other means
pct_work_home = percent of workers who work at home (commute is NA)
median_income = median household income
pct_health_ins = percent of civilian noninstitutionalized population with health insurance coverage
pct_private_health_ins = percent of civilian noninstitutionalized population with private health insurance
pct_public_health_ins = percent of civilian noninstitutionalized population with public health insurance
pct_no_health_ins = percent of civilian noninstitutionalized population without health insurance coverage
pct_impoverished = percent of families and individuals whose income in the past 12 months is below the poverty level

pop_under_5 = population under 5 years old
pop_5_9 = population between 5 and 9 years old
pop_10_14 = population 10 to 14 years old
pop_15_19 = population 15-19 years old
pop_20_24 = population 20-24 years old
pct_white = percent of total population white (either alone or in combination with other races)
pct_black = percent of total population black or african american
pct_native = percent "american indian" or alaska native
pct_asian = percent asian
pct_pacific_islander = percent native hawaiian or other pacific islander
pct_race_other = percent some other race
pct_hispanic_latinx = percent of total population hispanic or latinx

```{r ACS Data}
#Read in American Community Survey Social Characteristics for 2017 dataset
ACS_social <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/ACS_OK_VA_Social.csv")[-c(1),]%>%
  mutate(county = str_extract(`GEO.display-label`, ".+?(?=, )"),
         state = str_extract(`GEO.display-label`, "[^, ]+$")) %>%
  select(county, state, HC01_VC76, HC01_VC77, HC01_VC78, HC01_VC79, HC01_VC80, HC01_VC81, HC03_VC86, HC03_VC87, HC03_VC88, HC03_VC89, HC03_VC90, 
         HC03_VC91, HC03_VC92, HC03_VC95, HC03_VC96)

names(ACS_social) <- c("county", "state", "number_enrolled", "number_enrolled_prek", "number_enrolled_k", "number_enrolled_elementary", "number_enrolled_hs", "number_enrolled_college", "pct_under_9th", "pct_9th_12th", "pct_attained_hs", "pct_some_college", "pct_associates", "pct_bachelors", "pct_grad_degree", "pct_hs_or_higher", "pct_college_or_higher") 


#Read in American Community Survey Housing Characteristics for 2017 dataset
ACS_housing <-read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/ACS_OK_VA_Housing.csv")[-c(1),] %>%
  mutate(county = str_extract(`GEO.display-label`, ".+?(?=, )"),
         state = str_extract(`GEO.display-label`, "[^, ]+$")) %>%
  select(county, state, HC03_VC65, HC03_VC66, HC03_VC75, HC03_VC76, HC03_VC77, HC03_VC78, HC03_VC79, HC03_VC80)

names(ACS_housing) <- c("county", "state", "pct_owner_occupied_homes", "pct_renter_occupied_homes", "pct_moved_after_2014", "pct_moved_2010_2014", "pct_moved_2000_2009", "pct_moved_1990_1999", "pct_moved_1980_1989", "pct_moved_1979_earlier")


#Read in American Community Survey Economic Characteristics for 2017 dataset
ACS_economic <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/ACS_OK_VA_Economic.csv")[-c(1),] %>%
  mutate(county = str_extract(`GEO.display-label`, ".+?(?=, )"),
         state = str_extract(`GEO.display-label`, "[^, ]+$")) %>%
  select(county, state, HC03_VC28, HC03_VC29, HC03_VC30, HC03_VC31, HC03_VC32, HC03_VC33, HC01_VC36, 
         #HC03_VC75, HC03_VC76, HC03_VC77, HC03_VC78, HC03_VC79, HC03_VC80, HC03_VC81, HC03_VC82, HC03_VC83, HC03_VC84, 
         HC01_VC85, HC03_VC131, HC03_VC132, HC03_VC133, HC03_VC134, HC03_VC161)

names(ACS_economic) <- c("county", "state", "pct_commute_drive", "pct_commute_carpool", "pct_commute_public", "pct_commute_walk", "pct_commute_other", "pct_work_home", "commute_time_mean",
#"pct_income_under_10k", "pct_income_10k_15k", "pct_income_15k_25k", "pct_income_25k_35k", "pct_income_35k_50k", "pct_income_50k_75k", "pct_income_75k_100k", "pct_income_100k_150k", "pct_income_150k_200k", "pct_income_over_200k", 
                         "median_income", "pct_health_ins", "pct_private_health_ins", "pct_public_health_ins", "pct_no_health_ins", "pct_impoverished")


#Read in American Community Survey Demographic and Housing Characteristics for 2017 dataset
ACS_demographic <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/ACS_OK_VA_Demographic.csv")[-c(1),] %>%
  mutate(county = str_extract(`GEO.display-label`, ".+?(?=, )"),
         state = str_extract(`GEO.display-label`, "[^, ]+$")) %>%
  select(county, state, HC01_VC03, HC01_VC09, HC01_VC10, HC01_VC11, HC01_VC12, HC01_VC13, HC03_VC83, HC03_VC84, HC03_VC85, HC03_VC86,
         HC03_VC87, HC03_VC88, HC03_VC93)

names(ACS_demographic) <- c("county", "state", "population", "pop_under_5", "pop_5_9", "pop_10_14", "pop_15_19", "pop_20_24", "pct_white",
                            "pct_black", "pct_native", "pct_asian", "pct_pacific_islander", "pct_race_other", "pct_hispanic_latinx")


#merge all seperate ACS datasets with factors we want 
ACS <- ACS_demographic %>%
  full_join(ACS_economic, by=c("county", "state"))%>%
  full_join(ACS_housing, by=c("county", "state"))%>%
  full_join(ACS_social, by=c("county", "state"))%>%
  mutate(county = tools::toTitleCase(as.character(county)))%>%
  mutate(pct_enrolled_prek = (as.numeric(number_enrolled_prek)/as.numeric(pop_under_5))*100, pct_enrolled_elementary_middle = (as.numeric(number_enrolled_elementary)/(as.numeric(pop_5_9) + as.numeric(pop_10_14)))*100, pct_enrolled_hs = (as.numeric(number_enrolled_hs)/as.numeric(pop_15_19))*100, pct_less_than_hs = as.numeric(pct_under_9th) + as.numeric(pct_9th_12th))%>%
  select(-c(pop_under_5, pop_5_9, pop_10_14, pop_15_19, pop_20_24, number_enrolled, number_enrolled_prek, number_enrolled_k, number_enrolled_elementary, number_enrolled_hs, number_enrolled_college, pct_under_9th, pct_9th_12th))

va_ACS <- ACS %>%
  filter(state == "Virginia")

ok_ACS <- ACS %>%
  filter(state == "Oklahoma")

#removes now-unnecessary data from environment
rm(ACS, ACS_demographic, ACS_economic, ACS_housing, ACS_social)
```

## County Health Rankings Data

Explanation of factors:
fips = FIPS code
pct_fair_poor_health = percent of adults that report fair or poor health (as opposed to good or excellent)
teen_birth_rate = teen births/females ages 15-19 * 100,000
pcp_ratio = ratio of primary care physicians to population
pct_children_poverty = percent of children (under 18) living in poverty
income_ratio = ratio of household income at the 80th percentile to income at the 20th percentile
social_assoc_rate = social associations / population * 100,000

air_pollution = average daily PM2.5 (particulate matter)
water_violations = county affected by drinking water violation
severe_housing_problems = percentage of households with at least 1 of 4 housing problems: overcrowding, high housing costs, or lack of kitchen or plumbing facilities

pct_food_insecure = percent of population food insecure
median_income_white = median income of white households
median_income_black = median income of black households
median_income_hispanic = median income of hispanic households
segregation_black_white = index of residential segregation black/white
segregation_white_non = index of residential segregation non-white/white
pct_rural = percent of population in rural areas


```{r County Health}
#read first row of Virginia County Health data
first_row <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/2017_County_Health_Rankings_Virginia_Data-v2.xls", sheet = 4)
write.csv(first_row,'first_row.csv')

#combine first two rows for County Health
header <- sapply(read.csv("first_row.csv", nrow=2), paste, collapse="_")
header <- sapply(read.csv("first_row.csv", header=FALSE, nrow=2), paste, collapse="_")

#set combined first two rows as header, and select variables we want
county_health_ranked_measure_data <- read.csv("first_row.csv", header = FALSE, skip=3, col.names=header) %>% 
  select("X__2_State", "X__3_County","Poor.or.fair.health_..Fair.Poor", "X__45_Teen.Birth.Rate", "X__54_PCP.Ratio", "X__62_Preventable.Hosp..Rate",
         "X__88_..Unemployed", "Children.in.poverty_..Children.in.Poverty", "X__97_Income.Ratio", "X__104_Association.Rate", "Air.pollution...particulate.matter_Average.Daily.PM2.5",
         "Drinking.water.violations_Presence.of.violation", "X__114_..Severe.Housing.Problems")

names(county_health_ranked_measure_data) <- c("state", "county", "pct_fair_poor_health", "teen_birth_rate", "pcp_ratio", "preventable_hosp_rate", "pct_unemployed", "pct_children_poverty", "income_ratio", "social_assoc_rate", "air_pollution", "water_violations", "severe_housing_problems")


#read Virginia County Health additional measures
first_row <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/2017_County_Health_Rankings_Virginia_Data-v2.xls", sheet = 5)
write.csv(first_row,'first_row.csv')

#combine first two rows for Virginia additional data
header <- sapply(read.csv("first_row.csv", nrow=2), paste, collapse="_")
header <- sapply(read.csv("first_row.csv", header=FALSE, nrow=2), paste, collapse="_")

#set combined first two rows as header, and select variables we want
county_health_additional_measure_data <- read.csv("first_row.csv", header = FALSE, skip=3, col.names=header) %>% 
  select("X__2_State", "X__3_County", "X__20_..Food.Insecure", "X__37_Household.Income..white.alone.", "X__38_Household.Income..black.alone.", "X__39_Household.income..Hispanic.", "Residential.segregation...black.white_Segregation.index", "Residential.segregation...non.white.white_Segregation.Index", "X__65_..Rural")

names(county_health_additional_measure_data) <- c("state", "county", "pct_food_insecure", "median_income_white", "median_income_black", "median_income_hispanic", "segregation_black_white", "segregation_white_non", "pct_rural")

#combine ranked and additional data
va_county_health_rankings <- county_health_ranked_measure_data %>%
  right_join(county_health_additional_measure_data, by=c("state", "county"))%>%
  clean_county_names()


#read in Oklahoma County Health data
first_row <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/2017_County_Health_Rankings_Oklahoma_Data-v2.xls", sheet = 4)
write.csv(first_row,'first_row.csv')

#combines first two rows Oklahoma County Health data
header <- sapply(read.csv("first_row.csv", nrow=2), paste, collapse="_")
header <- sapply(read.csv("first_row.csv", header=FALSE, nrow=2), paste, collapse="_")

#set combined first two rows as header, and select variables we want
county_health_ranked_measure_data <- read.csv("first_row.csv", header = FALSE, skip=3, col.names=header) %>% 
  select("X__2_State", "X__3_County","Poor.or.fair.health_..Fair.Poor", "X__45_Teen.Birth.Rate", "X__54_PCP.Ratio", "X__62_Preventable.Hosp..Rate",
         "X__88_..Unemployed", "Children.in.poverty_..Children.in.Poverty", "X__97_Income.Ratio", "X__104_Association.Rate", "Air.pollution...particulate.matter_Average.Daily.PM2.5",
         "Drinking.water.violations_Presence.of.violation", "X__114_..Severe.Housing.Problems")

names(county_health_ranked_measure_data) <- c("state", "county", "pct_fair_poor_health", "teen_birth_rate", "pcp_ratio", "preventable_hosp_rate", "pct_unemployed", "pct_children_poverty", "income_ratio", "social_assoc_rate", "air_pollution", "water_violations", "severe_housing_problems")


#read Oklahoma County Health additional measures
first_row <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/2017_County_Health_Rankings_Oklahoma_Data-v2.xls", sheet = 5)
write.csv(first_row,'first_row.csv')

#combine first two rows of Oklahoma County Health additional measures
header <- sapply(read.csv("first_row.csv", nrow=2), paste, collapse="_")
header <- sapply(read.csv("first_row.csv", header=FALSE, nrow=2), paste, collapse="_")

#set combined first two rows as header, and select variables we want
county_health_additional_measure_data <- read.csv("first_row.csv", header = FALSE, skip=3, col.names=header) %>% 
  select("X__2_State", "X__3_County", "X__20_..Food.Insecure", "X__37_Household.Income..white.alone.", "X__38_Household.Income..black.alone.", "X__39_Household.income..Hispanic.", "Residential.segregation...black.white_Segregation.index", "Residential.segregation...non.white.white_Segregation.Index", "X__65_..Rural")

names(county_health_additional_measure_data) <- c("state", "county", "pct_food_insecure", "median_income_white", "median_income_black", "median_income_hispanic", "segregation_black_white", "segregation_white_non", "pct_rural")


#combine Oklahoma ranked and additional measures
ok_county_health_rankings <- county_health_ranked_measure_data %>%
  right_join(county_health_additional_measure_data, by=c("state", "county"))%>%
  clean_county_names()

#removes now-unnecessary data from environment
rm(county_health_additional_measure_data, county_health_ranked_measure_data, first_row, header)
```

## Library Data (ended up not using)

Explanation of factors:
county = county in which the headquarters of the administrative entity is physically located
lib_lsa_population = Population of the Legal Service Area
lib_hours_open = Total annual public service hours for all service outlets
lib_circulation = total annual circulation transactions
lib_kids_circulation = Total annual circulation (including renewals) of all children's materials in all formats to all users
lib_program = total library programs
lib_kids_program = total children's programs
lib_ya_program = total young adult programs
lib_prog_atten = total audience at all library programs
lib_kids_prog_atten = total audience at all children's programs
lib_ya_prog_atten = total audience at all young adult programs

```{r}
#read in library data 
library_data <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/va_library_data.csv") %>%
  select("STABR", "CNTY", "POPU_LSA", "HRS_OPEN", "VISITS", "REGBOR", "TOTPRO", "TOTATTEN")

names(library_data) <- c("state_abr", "county", "lib_lsa_population", "lib_hours_open", "lib_yearly_visits", "lib_registered_users", "lib_program", "lib_prog_atten")

#combine library information by county, and extract meaningful parameters
va_ok_library <- library_data %>%
  filter(state_abr == "VA" | state_abr == "OK")%>%
  clean_county_names()%>%
  group_by(county, state_abr)%>%
  summarise(lib_lsa_population = sum(lib_lsa_population), lib_hours_open = sum(lib_hours_open), lib_yearly_visits = sum(lib_yearly_visits), lib_registered_users = sum(lib_registered_users), lib_program = sum(lib_program), lib_prog_atten = sum(lib_prog_atten))%>%
  mutate(pct_lib_program_atten = lib_prog_atten/lib_lsa_population, lib_visit_rate = lib_yearly_visits/lib_lsa_population, pct_lib_patrons = lib_registered_users/lib_lsa_population, lib_hours_per_person = lib_hours_open/lib_lsa_population)%>%
  select(-c(lib_lsa_population, lib_hours_open, lib_yearly_visits, lib_registered_users, lib_program, lib_prog_atten))
  
#select library data for Virginia
va_library <- va_ok_library %>%
  filter(state_abr == "VA")%>%
  select(-c(state_abr))%>%
  ungroup(county)

#King and Queen and King William counties share a library system with Hanover (in row 39)
#Stafford shares a library system with Spotsylvania (in row 79)
king_queen_lib <- va_library[39,]%>%
  mutate(county = "King and Queen County")
king_william_lib <- va_library[39,]%>%
  mutate(county = "King William County")
stafford_lib <- va_library[79,] %>%
  mutate(county = "Stafford County")

#add library data for King and Queen, King William, and Stafford Counties
va_library <- rbind(va_library, king_queen_lib, king_william_lib, stafford_lib)

#select library data for Oklahoma
ok_library <- va_ok_library %>%
  filter(state_abr == "OK")%>%
  select(-c(state_abr))%>%
  ungroup(county)

#removes now-unnecessary data from environment
rm(library_data, va_ok_library, king_queen_lib, king_william_lib, stafford_lib)
```

## Migration Data

Explanation of Factors:

pct_stayed = Total number of those who did not move or moved within county, divided by population

```{r}
#read Migration Data for Virginia
first_row <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/county-to-county-2012-2016-previous-residence-sort.xlsx", sheet = "Virginia", skip = 1)
write.csv(first_row,'first_row.csv')

#combine first two rows of Virginia Migration Data
header <- sapply(read.csv("first_row.csv", nrow=2), paste, collapse="_")
header <- sapply(read.csv("first_row.csv", header=FALSE, nrow=2), paste, collapse="_")

#set first two rows as header, and select variables we want
va_migration <- read.csv("first_row.csv", header = FALSE, skip=3, col.names=header) %>%
  select("State.U.S..Island.Area.Foreign.Region.of.Residence.1.Year.Ago_NA", "County.of.Residence.1.Year.Ago_NA", "County.of.Residence.1.Year.Ago__1_Population.1.Year.and.Over", "X__15_Nonmovers", "X__19_Movers.within.Same.County") %>% 
  mutate(County.of.Residence.1.Year.Ago_NA = tools::toTitleCase(as.character(County.of.Residence.1.Year.Ago_NA)))

names(va_migration) <- c("state", "county", "population", "nonmovers", "movers_within_county") 

#remove duplicate rows
va_migration <- unique(va_migration)
#remove row of all NAs
va_migration <- va_migration[!(rowSums(is.na(va_migration))==NCOL(va_migration)),]

#figure out what percentage of population stayed within county
va_migration <- va_migration %>%
  mutate(pct_stayed = (nonmovers+movers_within_county)/population)%>%
  mutate(pct_movers_within_county = movers_within_county/(population-nonmovers))%>%
  select(-c(population, nonmovers, movers_within_county))


#read Migration Data for Oklahoma
first_row <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/county-to-county-2012-2016-previous-residence-sort.xlsx", sheet = "Oklahoma", skip = 1)
write.csv(first_row,'first_row.csv')

#combine first two rows of Oklahoma Migration Data
header <- sapply(read.csv("first_row.csv", nrow=2), paste, collapse="_")
header <- sapply(read.csv("first_row.csv", header=FALSE, nrow=2), paste, collapse="_")

#set first two rows as header, and select variables we want
ok_migration <- read.csv("first_row.csv", header = FALSE, skip=3, col.names=header) %>%
  select("State.U.S..Island.Area.Foreign.Region.of.Residence.1.Year.Ago_NA", "County.of.Residence.1.Year.Ago_NA", "County.of.Residence.1.Year.Ago__1_Population.1.Year.and.Over", "X__15_Nonmovers", "X__19_Movers.within.Same.County") %>% 
  mutate(County.of.Residence.1.Year.Ago_NA = tools::toTitleCase(as.character(County.of.Residence.1.Year.Ago_NA)))

names(ok_migration) <- c("state", "county", "population", "nonmovers", "movers_within_county") 

#remove duplicate rows
ok_migration <- unique(ok_migration)
#remove row of all NAs
ok_migration <- ok_migration[!(rowSums(is.na(ok_migration))==NCOL(ok_migration)),]

#figure out what percentage of population stayed within county
ok_migration <- ok_migration %>%
  mutate(pct_stayed = (nonmovers+movers_within_county)/population)%>%
  mutate(pct_movers_within_county = movers_within_county/(population-nonmovers))%>%
  select(-c(population, nonmovers, movers_within_county))

#removes now-unnecessary data from environment
rm(first_row, header)
```


## Kids Count Data

Explanation of factors:

pct_child_no_ins = Percent of all children in county who do not have health insurance
child_abuse_rate = rate of child abuse per 1,000 children
pct_foster_1-5 = percent of children ages 1 through 5 who are in foster care
pct_foster_11-15 = percent of children ages 11 through 15 who are in foster care
pct_foster_16-18 = percent of children ages 16 through 18 who are in foster care
pct_foster_6-10 = percent of children ages 6 through 10 who are in foster care
pct_foster_under_1 = percent of infants (under 1) who are in foster care
foster_rate = number of children in foster care per 1000 children

```{r}
#read Kids Count data for Virginia
va_child_ins <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_children-uninsured_virginia_2016.xlsx") %>%
  select(Location, percent)
names(va_child_ins) <- c("county", "pct_child_no_ins")
va_child_food <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_child-food_insecurity_Virginia_2016.xlsx")%>%
  select(Location, Percent)
names(va_child_food) <- c("county", "pct_child_food_insecure")
va_child_abuse <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_child-abuse_virginia_2017.xlsx")
names(va_child_abuse) <- c("county", "child_abuse_rate")

#join all Kids Count data for Virginia
va_kids_count <- va_child_ins %>%
  full_join(va_child_food, by = c("county")) %>%
  full_join(va_child_abuse, by = c("county")) %>%
  filter(county != "Data Type") %>%
  clean_county_names()

#removes now-unnecessary data from environment
rm(va_child_ins, va_child_food, va_child_abuse)


#read Kids Count Data for Oklahoma
ok_child_ins <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_children-uninsured_oklahoma_2016.xlsx") %>%
  select(location, percent)
names(ok_child_ins) <- c("county", "pct_child_no_ins")
ok_child_food <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_child-food_insecurity_oklahoma_2016.xlsx")%>%
  select(Location, Percent)
names(ok_child_food) <- c("county", "pct_child_food_insecure")
ok_child_abuse <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_child-abuse_oklahoma_2017.xlsx")
names(ok_child_abuse) <- c("county", "child_abuse_rate")
ok_foster <- read_excel("/home/jk9ra/ari_social_media/data/working/County_Level/Kids Count/KC_foster-care_oklahoma_2017.xlsx")%>%
  select(Location, rate_per_1000)
names(ok_foster) <- c("county", "foster_rate")

#join all Kids Count data for Oklahoma
ok_kids_count <- ok_child_ins %>%
  full_join(ok_child_food, by = c("county")) %>%
  full_join(ok_child_abuse, by = c("county")) %>% 
  filter(county != "Age group")

#removes now-unnecessary data from environment
rm(ok_child_ins, ok_child_food, ok_child_abuse, ok_foster)
```


## Death data (ended up not using)

Explanation of factors:

death_rate =  total deaths in county in 2016/population * 1000

```{r}
#read death (injury in this case is code for death) data for Virginia
va_death <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/Deaths_2016_Data_Request_Dataset.csv") %>%
  dcast(Locality + population ~ INDICATOR, value.var = "deaths") %>%
  mutate(Locality = ifelse(Locality %in% va_city_county_overlap, str_c(as.character(Locality), "County", sep = " "),
                                ifelse(str_detect(Locality, "City") | str_detect(Locality, "County"), as.character(Locality),
                                       ifelse(Locality %in% va_independent_cities, str_c(as.character(Locality), "City", sep = " "), str_c(as.character(Locality), "County", sep = " ")))))%>%
  mutate(death_rate = (ALLINJURY/population)*1000)%>%
  select(-c(population, ALLINJURY, FALLINJURY, FIREARMSINJURY, HOMICIDEINJURY, MVTINJURY, OVERDOSEINJURY, POISONINGINJURY, SUICIDEINJURY, TBIINJURY))

names(va_death) <- c("county", "death_rate")


#read death (injury in this case is code for death) data for Oklahoma
ok_death <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/Deaths_2016_OK.csv")%>%
  mutate(death_rate = (`All Injury`/Population)*1000)%>%
  select(-c(Year))%>%
  mutate(Locality = str_c(as.character(Locality),"County", sep = " "))%>%
  select(-c(`All Injury`, Fall, Firearms, Homicide, `Motor Vehicle`, Poisoning, Overdose, Suicide, TBI, Population))

names(ok_death) <- c("county", "death_rate")
```

## Voting Information

Explanation of factors:

total_votes = number of ballots cast in most recent gubenatorial election (2017 for VA, 2018 for OK)
total_registered = number of registered voters in county

```{r}
#read voting information for Virginia
va_voting <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/governor_race_virginia_totals_2017.csv")%>%
  mutate(county = tools::toTitleCase(as.character(county_name)))%>%
  clean_county_names()%>%
  select(-c(county_name))

#read voting information for Oklahoma
ok_voting <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/governor_race_oklahoma_totals_2018.csv")%>%
  mutate(county = tools::toTitleCase(as.character(county_name)))%>%
  clean_county_names()%>%
  select(-c(county_name))
```

## Fast food, churches, and liquor (ended up not using)

Explanation of factors:

num_fast_food = number of fast food establishments in county
num_liquor_stores = number of liquor stores in county
num_places of worship = number of places of worship (churches, mosques, synagogues, etc) in county

```{r}
#read in data
worship_fast_food <- read_csv("/home/jk9ra/ari_social_media/data/working/County_Level/places_of_worship_fast_food_liquor (1).csv") %>%
  group_by(county_name, type)%>%
  summarize(number = n())%>%
  dcast(county_name ~ type, value.var = "number")

names(worship_fast_food) <- c("county", "num_fast_food", "num_liquor_stores", "num_places_of_worship")

worship_fast_food <- clean_county_names(worship_fast_food)
```



## Combine all Virginia and Oklahoma data, respectively

```{r}
#combine all Virginia data
va_indicators <- va_ACS %>%
  full_join(va_county_health_rankings, by = c("state", "county"))%>%
  full_join(va_kids_count, by = c("county"))%>%
  full_join(va_migration, by = c("state", "county"))%>%
  full_join(va_library, by = c("county"))%>%
  full_join(va_death, by = c("county"))%>%
  full_join(va_voting, by = c("county"))%>%
  select(-c("state"))

#use population data to calculate percent registered and percent voted
va_indicators <- cbind(va_indicators[,1], sapply(va_indicators[,-1], as.numeric))%>%
  mutate(pct_registered = total_registered/population, pct_voted = total_votes/population, pct_reg_voted = total_votes/total_registered)%>%
  select(-c("total_registered", "total_votes"))

#combine all Oklahoma data
ok_indicators <- ok_ACS %>%
  full_join(ok_county_health_rankings, by = c("state", "county"))%>%
  full_join(ok_kids_count, by = c("county"))%>%
  full_join(ok_migration, by = c("state", "county"))%>%
  full_join(ok_library, by = c("county"))%>%
  full_join(ok_death, by = c("county"))%>%
  full_join(ok_voting, by = c("county"))%>%
  select(-c(state))

#use population data to calculate percent registered and percent voted
ok_indicators <- cbind(ok_indicators[,1], sapply(ok_indicators[,-1], as.numeric))%>%
  mutate(pct_registered = total_registered/population, pct_voted = total_votes/population, pct_reg_voted = total_votes/total_registered)%>%
  select(-c("total_registered", "total_votes"))

#removes now-unnecessary data from environment
rm(va_ACS, va_county_health_rankings, va_kids_count, va_migration, va_library, va_death, va_voting, ok_ACS, ok_county_health_rankings, ok_kids_count, ok_migration, ok_library, ok_death, ok_voting, va_independent_cities, va_city_county_overlap)


#isolates data for Virginia counties of interest
va_counties <- va_indicators %>%
  filter(county %in% c("Stafford County", "Spotsylvania County", "Hanover County", "King William County", "King and Queen County", "Essex County", "King George County", "Caroline County"))%>%
  inner_join(worship_fast_food, by = "county")%>%
  mutate(church_rate = (num_places_of_worship/population)*100000, fast_food_rate = (num_fast_food/population)*100000, liquor_store_rate = (num_liquor_stores/population)*100000)%>%
  select(-c("num_places_of_worship", "num_fast_food", "num_liquor_stores"))

#isolates data for Oklahoma counties of interest
ok_counties <- ok_indicators %>%
  filter(county %in% c("Caddo County", "Grady County", "Stephens County", "Cotton County", "Tillman County", "Kiowa County", "Comanche County"))%>%
  inner_join(worship_fast_food, by = "county")%>%
  mutate(church_rate = (num_places_of_worship/population)*100000, fast_food_rate = (num_fast_food/population)*100000, liquor_store_rate = (num_liquor_stores/population)*100000)%>%
  select(-c("num_places_of_worship", "num_fast_food", "num_liquor_stores"))
```